---
title: "Final Project-Mapping"
author: "Peace Corp"
date: "5/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plyr)
library(dplyr)
library(readr)
library(purrr)
library(httr)
library(viridis)
library(lubridate)
library(hms)
library(janitor)
library(sf)
library(tidycensus)
library(patchwork)
library(stringr)
library(ggplot2)
#read and load DC crimes data
crimes <-
  list.files(path = "data/",
             pattern = "*.csv", 
             full.names = T) %>% 
  map_df(~read_csv(., col_types = cols("CCN"= col_double()))) 
```


```{r}
#change case of the variables
change_case<-function(x){
  return(gsub("\\b([A-Z])([A-Z]+)", "\\U\\1\\L\\2", x, perl=TRUE))
}
#change case and split date variable to weekdays, month and year, label property offense 
crimes<-as.tibble(crimes)%>%
  clean_names()%>%
  dplyr::mutate(shift=change_case(shift),
                method=change_case(method),
                offense=change_case(offense),
                block=change_case(block),
                bid=change_case(bid),
                weekday=lubridate::wday((start_date), label=TRUE),
                month=lubridate::month((start_date),label=TRUE),
                year=lubridate::year(report_dat),
                occur_date=date(start_date),
                occur_time=hms::as_hms(ymd_hms(start_date)),
               # time_group=lubridate::hour(occur_time) %/% 6),
               #   cut(as.numeric(occur_time),
               #  breaks=c(0,6*60,12*60,18*60,23*60+59),
              #  labels=c("00-06","06-12","12-18","18-00"),include.lowest = TRUE),
                crimetype=ifelse(offense%in%c("robbery","theft","burglary"), "property", "non-property")
  )
#identify duplication incidents by CCN
crimes%>%
  group_by(ccn)%>%
  dplyr::mutate(count=n())%>%
  filter(count>1)
#remove duplicate rows 
crimes_clean<-filter(distinct(crimes, ccn, .keep_all = TRUE))
#check 
crimes_clean%>%
  group_by(ccn)%>%
  dplyr::mutate(count=n())%>%
  filter(count>1)
crimes_clean%>%
  group_by(offense)%>%
  dplyr::summarize(count=n())%>%
  arrange(desc(count))
```

```{r}
#load the census tracts of DC 
unzip(zipfile="data/tl_2017_11_tract.zip",
      exdir = "data")
#read shapefile into sf
dc<-st_read("data/tl_2017_11_tract.shp")%>%
  select(GEOID, geometry)%>%
  clean_names()%>%
  st_transform(crs=4326)
crimes_sf<-st_as_sf(crimes_clean, coords = c("longitude", "latitude"),crs=4326)
```

#The distribution of "Theft F/Auto" incidence in District of Clombia 2017-2021
```{r}
#count the property-related "theft" crimes and perform a spatial join
dc_merged_agg<-st_join(dc, crimes_sf, join=st_intersects)%>%
  filter(offense=="Theft F/Auto")%>%
  group_by(geoid)%>%
  dplyr::summarize(count = n())
#create choropleth of the property-related "theft" crimes
count1<-dc_merged_agg%>%
 
  ggplot()+
  geom_sf(aes(fill=count), color = "white", size = 0.1)+
  scale_fill_viridis(option = "D")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.6))

count1
```

#The distribution of "Theft/Other" incidence in the District of Clombia 2017-2021
```{r}
#count the property-related "theft" crimes and perform a spatial join
dc_merged_agg<-st_join(dc, crimes_sf, join=st_intersects)%>%
  filter(offense=="Theft/Other")%>%
  group_by(geoid)%>%
  dplyr::summarize(count = n())
#create choropleth of the count of property-related "theft" crimes
count2<-dc_merged_agg%>%
 
  ggplot()+
  geom_sf(aes(fill=count), color = "white", size = 0.1)+
  scale_fill_viridis(option = "D")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.6))

count2
```


#The distribution of "Robbery" incidence in District of Clombia 2017-2021
```{r}
#count the property-related "Robbery" crimes and perform a spatial join
dc_merged_agg<-st_join(dc, crimes_sf, join=st_intersects)%>%
  filter(offense=="Robbery")%>%
  group_by(geoid)%>%
  dplyr::summarize(count = n())
#create choropleth of the count of property-related "Robbery" 
count3<-dc_merged_agg%>%
  ggplot()+
  geom_sf(aes(fill=count), color = "white", size = 0.1)+
  scale_fill_viridis(option = "D")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.6))
count3
```

#The distribution of "Burglary" incidence in District of Clombia 2017-2021
```{r}
#count the property-related "Burglary" crimes and perform a spatial join
dc_merged_agg<-st_join(dc, crimes_sf, join=st_intersects)%>%
  filter(offense=="Burglary")%>%
  group_by(geoid)%>%
  dplyr::summarize(count = n())

#create choropleth of the count of property-related "Burglary" 
count4<-dc_merged_agg%>%
  ggplot()+
  geom_sf(aes(fill=count), color = "white", size = 0.1)+
  scale_fill_viridis(option = "D")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.6))

count4
```

#Combining the mapping output

```{r}
library(ggpubr)
p=ggarrange(count1,count2,count3,count4,labels=c("Theft F/Auto 2017-2021","Theft/Other 2017-2021","Robbery 2017-2021","Burglary 2017-2021",ncol=4,
                                   common.legend=TRUE,legend="none"),hjust = -1,vjust =1.1,font.label =list(size = 9)) 
p
```

#The Spatial geographic distribution of Police Stations in District of Columbia 
```{r}
#Identify the Spatial geographic distribution of Police Stations in District of Columbia

police_stations<- st_read("/Users/yiwenwang/Desktop/Data\ Skills/R/Final\ Project-R/data/Police_Stations/Police_Stations.shp")

DC_tracks <- st_read("/Users/yiwenwang/Desktop/Data\ Skills/R/Final\ Project-R/data/tl_2017_11_tract/tl_2017_11_tract.shp")

```

#Count the number of crimes and perform a spatial join
```{r}
dc_merged_agg<-st_join(dc, crimes_sf, join=st_intersects)%>%
  group_by(geoid)%>%
  dplyr::summarize(count = n())

count<-dc_merged_agg%>%
  ggplot()+
  geom_sf(aes(fill=count), color = "white", size = 0.1)+
  scale_fill_viridis(option = "D")+
  geom_sf(data =police_stations,color="red")+theme_void()

count
```


